---
title: "HubGrub: Grubbing for information in the Hubs"
author:
- name: Kayla Interdonato
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
- name: Martin Morgan
  affiliation: Roswell Park Comprehensive Cancer Center, Buffalo, NY
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        toc: true
        toc_float: true
package: HubGrub
vignette: >
  %\VignetteIndexEntry{HubGrub: Grubbing for information in the Hubs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEndcoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Introduction

`HubGrub` is a package that provides users with functionality to discover data
in the Bioconductor Hub structures. The package helps the user discover what
type of data is available for retrieval and can display the metadata of that
data. There is also functionality to retrieve the data and import it without
needing to download the entire file. We will demonstrate all the functionality
of the package in the following vignette.

# Installation
Install the most recent version from Bioconductor:

```{r bioconductor, eval = FALSE}
if(!requireNamesapce("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("HubGrub")
``` 

The development version is also available for install from GitHub:

```{r github, eval = FALSE}
BiocManager::install("Kayla-Morrell/HubGrub")
```

Then load `HubGrub`:

```{r load, message = FALSE}
library(HubGrub)
```

# HubGrub

## Discover data in Hubs

The first step when looking into the hubs is to figure out what type of data is
available. Once a hub object is created, either from AnnotationHub or
ExperimentHub, then the user can use `discoverData()` to see how many resources
are available for a specific data type. For example, say we want to find out how
many bigwig files are in Bioconductor's AnnotationHub.

```{r discover}
library(AnnotationHub)
ah = AnnotationHub()
discoverData(ah, "BigWigFile")
```

## Info on data in Hubs

Once we figure out which files we are interested in we can use `dataInfo()` to
see the metadata information on these files. This is where the user can narrow
down their search to specific files they are interested in retrieving.

```{r info}
tbl <- dataInfo(ah, "BigWigFile")
tbl
```

A user maybe intersted in specifically human files of a specific cell type. Then
they can use `dplyr` functionality on the table to create the subset of files
they are interested in. 

```{r subset}
tbl <- tbl |> 
    dplyr::filter(species == "Homo sapiens") |> 
    dplyr::select(ID, title)
tail(tbl)
```

## Importing the files

With a subset of files that the user is interested in, `importData()` will read
in a part of the file without having to download the entire resource. These
files can be quite large, so importing just a portion of a file can be
beneficial to users.

In order to use this function to it's fullest potential, the user should define
a range of data that they are interested in. This range of data should be in the
form of a `GRanges` object.

```{r import}
library(GenomicRanges)
which <- GRanges(c("chr2", "chr2"), IRanges(c(1, 300), c(400, 1000000)))
importData(ah, "AH49544", which)
```

## Session Information

```{r session}
sessionInfo()
```
